Modeling Intervention Strategies for United States TB Control

### 2014-02-15 in_progress/presentations/paper/draft.tex
Modifying discussion to discuss plateauing effect
working on draft. Edited Abstract and starting Introduction

\section{Abstract}
An extended epidemiological model for Tuberculosis in the US was developed,
based on an existing epidemiological model developed in 2012 by the Centers for
Disease Control and Prevention.  Economic parameters to track the cost burden of
TB for the US health care system were added.  The sensitivity of the extended
model to changes in parameters was analyzed, and the percentage of foreign-born
immigrants with LTBI appears to be a highly influential parameter in the model.
Intervention strategies to reduce LTBI among foreign-born arrivals were
evaluated, and estimates of the cost per case averted were found. Additionally,
a population-level, stochastic, agent-based model was built to provide
statistical verification of the results of the deterministic model and to
demonstrate feasibility of agent-based modeling even on such large population
scales.

\section{Introduction}
Epidemiological models allow public health professionals to predict and analyze
disease dynamics and intervention effectiveness. The most common examples of
such models are compartmental differential equation models, in which the
population is split between several possible health states, with flow between
each state given according to deterministic differential equations. In 2012, 
Hill, Becerra, and Castro implemented a compartmental
differential equaiton model of tuberculosis (TB) in the United States (US).
Their model utilized five health states and two subpopulations, US-born (USB)
and foreign-born (FB) for a total of 10 compartments. They used this model to
evaluate several possible intervention strategies, and ultimately conclude that
though increasing LTBI treatment was a good intervention strategy, the US was
unlikely to meet their stated goal of elimination of TB in the US by 2100. In
this work, the Hill model was extended in several key ways. First, additional
tracking capabilities were added to the Hill model, such that it can now report
further granularity in the disease dynamics. Further, economic components were
added to the model in order to project the US health care system (HCS) costs due
to TB given our current policy as well as for various interventions.

@@

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.25]{figures/HillModelFlowChart.png}
  \end{center}
  \caption{Schematic of the Hill Model.}
  \label{fig:hillModelSchematic}
\end{figure}

@@

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|} \hline
       & red5    & red10    & red25    & red50    \\ \hline
  2000 & 0.00    & 0.00     & 0.00     & 0.00     \\ 
  2025 & 1115.12 & 2231.70  & 5592.43  & 11227.57 \\ 
  2050 & 3406.41 & 6821.27  & 17117.18 & 34447.71 \\ 
  2075 & 4976.77 & 9967.15  & 25021.33 & 50389.58 \\ 
  2100 & 5942.02 & 11900.90 & 29880.03 & 60189.49 \\ 
   \hline
\end{tabular}
\caption{Cases of TB Averted by Reducing Incoming LTBI by various percentages
         (Cure=\$800)} 
\label{tab:caRed}
\end{table}

>>>>

{\huge \color{red} Check if first person is better.}

\section{Abstract}
A deterministic, compartmental epidemiological and economic model for US
Tuberculosis was developed, extending a model developed in 2012 by the Centers
for Disease Control and Prevention. Sensitivity analysis of this model reveals
that the percentage of foreign-born immigrants with LTBI is a highly influential
parameter. Various intervention strategies, particularly those that reduced LTBI
among foreign-born arrivals, were evaluated, and cost per case averted estimates
were found.  Additionally, a population-level, stochastic, agent-based model was
built. This model demonstrated the feasibility of using agent-based modeling in
the context of minimal population heterogeneity and provided statistical
characterization of the results of the compartmental model. Specifically, it
revealed that final population sizes were normal, with the compartmental model
sampling the mean of these distributions.

\section{Introduction}
Epidemiological models allow public health professionals to predict and analyze
disease dynamics and intervention effectiveness. The most common examples of
such models are detereministic, compartmental, differential equation models. In
these models, the population is split between several possible health states, or
compartments, and a system of differential equations governs population flow
between the various compartments. In 2012, Hill, Becerra, and Castro implemented
a compartmental differential equaiton model of tuberculosis (TB) in the United
States (US).  Their model characterized five health states across two
subpopulations, US-born (USB) and foreign-born (FB), for a total of 10
compartments. They used this model to evaluate several possible intervention
strategies and ultimately concluded that though increasing LTBI treatment was an
effective intervention strategy, the US was unlikely to meet their stated goal
of elimination of TB in the US by 2100. Here, we present an extended Hill model
in which additional
tracking capabilities were added to the Hill model, such that it can now report
further granularity in the disease dynamics. Further, economic components were
added to the model in order to project the US health care system (HCS) costs due
to TB given our current policy as well as for various interventions.

@@

{\huge \color{red}TODO: Update this to the one I made for my talk}
\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.25]{figures/HillModelFlowChart.png}
  \end{center}
  \caption{Schematic of the Hill Model.}
  \label{fig:hillModelSchematic}
\end{figure}

@@

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Mon Oct 14 00:36:15 2013
\begin{table}
\centering
\begin{tabular}{|r|cccc|} \hline
       & red5    & red10    & red25    & red50    \\ \hline
  %2000 & 0.00    & 0.00     & 0.00     & 0.00     \\ 
  2025 & 1115.12 & 2231.70  & 5592.43  & 11227.57 \\ 
  2050 & 3406.41 & 6821.27  & 17117.18 & 34447.71 \\ 
  2075 & 4976.77 & 9967.15  & 25021.33 & 50389.58 \\ 
  2100 & 5942.02 & 11900.90 & 29880.03 & 60189.49 \\ 
   \hline
\end{tabular}
\caption{Cases of TB Averted by Reducing Incoming LTBI by various percentages
         (Cure=\$800)} 
\label{tab:caRed}
\end{table}
